
Practical Machine Learning Project:  
====================================
  
  
**Objective**
----------------
  
The goal of the project is to build a Model to predict how well individuals do barbell lifts using as input data from accelerometers on the belt, forearm, arm, and dumbbell of six participants. 
  
 
 
**Data**
---------------
  
The Weight Lifting Exercises Dataset for this project come from this source:http://groupware.les.inf.puc-rio.br/har.

The training data for this project are available here: 

  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


For detailed information see this paper:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3PlcqRRv0
  
  
**Steps**
------------
 
   
0. Reading data and creating of two datasets for training and testing model


```{r, eval=FALSE}
pml_Training <- read.csv("pml-training.csv")
library(caret)
intrain <- createDataPartition(y=pml_Training$classe, p=0.75, list=FALSE)
pmlTrain <- pml_Training[intrain,]
pmlTest <- pml_Training[-intrain,]
```


1. Predictors selection: choose the best set of variables for the fit model

First, it was removed any variable which contained missing values.  
Second, the descriptive variables that do not influence the final result were discarded: user name,the first three columns which recording time, new_window and num_window.  
Third, a visual inspection of the variables was performed using boxplots. In general, the predictive power of each variable is weak separately.

```{r, eval=FALSE}
vChoice <- c(8:11,37:49,60:68,84:86,102,113:124,140,151:159,160)
pmlTrain_1 <- pmlTrain[,vChoice]
pmlTest_1 <- pmlTest[,vChoice]
```

Examples of boxplots to see the data range of each variable associated with each
class (A,B,C,D,E)

```{r, echo=FALSE}
par(mfrow = c(1, 2))
boxplot(pmlTrain[,9]~classe,data=pmlTrain, ylab= "roll_belt")
boxplot(pmlTrain[,10]~classe,data=pmlTrain, ylab= "pitch_belt")
par(mfrow = c(1, 1))
```


2. Model Selection and Cross validation

 The model was fitted with a gradient boosted machine (gbm) using the selected predictors and a K-fold cross-validation without repetition. The 'gbm' method is adecuated because the predictors are weak and this method allows taking these variables, and average them together with weights, in order to get a stronger predictor.

```{r, eval=FALSE}
library(gbm)
fitControl <- trainControl(method = "cv", number = 10,repeats = 10)
pmlFit <-train(classe ~.,method="gbm",trControl=fitControl,verbose=FALSE,data=pmlTrain_1)
pmlFit
```


3. Test of model with testing set 

The model was tested with testing data set (pmlTest_1)

```{r, eval=FALSE}
pml_pred <- predict(pmlFit,pmlTest_1)
confusionMatrix(pmlTest_1$classe,pml_pred)
```


4. Validation of Model with 20 test cases

Finally, the machine learning algorithm was used to predict 20 different test cases. The results were very satisfactory with 20 out of 20 correct. 

```{r, eval=FALSE}
pmlTest3 <- read.csv("pml-testing.csv")
pmlTest3 <- pmlTest3[,vChoice]
pred <- predict(pmlFit,pmlTest3)
```

The confusion Matrix between observed and predicted data shows an 100% accuracy
along with a confidence interval betwen 0.83 and 1.  
For this test the model shows a 100% sensitivity and 100% Specificity. 


Confusion Matrix and Statistics

          Reference  
Prediction A B C D E  
         A 7 0 0 0 0  
         B 0 8 0 0 0  
         C 0 0 1 0 0  
         D 0 0 0 1 0  
         E 0 0 0 0 3  

Overall Statistics
                                     
               Accuracy : 1          
                 95% CI : (0.8316, 1)  
    No Information Rate : 0.4        
    P-Value [Acc > NIR] : 1.1e-08    
                                     
                  Kappa : 1          
 Mcnemar's Test P-Value : NA         

Statistics by Class:

                     Class: A Class: B Class: C Class: D Class: E  
Sensitivity              1.00      1.0     1.00     1.00     1.00  
Specificity              1.00      1.0     1.00     1.00     1.00  
Pos Pred Value           1.00      1.0     1.00     1.00     1.00  
Neg Pred Value           1.00      1.0     1.00     1.00     1.00  
Prevalence               0.35      0.4     0.05     0.05     0.15  
Detection Rate           0.35      0.4     0.05     0.05     0.15  
Detection Prevalence     0.35      0.4     0.05     0.05     0.15  
Balanced Accuracy        1.00      1.0     1.00     1.00     1.00  

```{r, eval=FALSE}
observed <- c("B","A","B","A","A","E","D","B","A","A","B","C",
              "B","A","E","E","A","B","B","B")
matrix <- confusionMatrix(observed,pred)
```

